---
title: "pokemonAnalysisForFun"
author: "J"
date: "2024-12-27"
output: html_document
---



12/31 = What to improve on: Making things look pretty, need to review notes more



Just for fun analysis of Pokemon :)

```{r message=FALSE, warning=FALSE}
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable     # converting tables to HTML
)
```

Import data

```{r}
pokemon <- import("Pokemon.csv")
```

Exploratory Data Analysis

preview of the data
```{r}
skim(pokemon)
```

taking a look at the structure/type
```{r}
str(pokemon)
```

Cleaning Up Names (making it into lower case and switching space for underscore)

```{r}
pokemon <- pokemon %>% janitor::clean_names()

str(pokemon)
```

~Question how many pokemon have only 1 Type vs 2 Types?

Using Skimr, 386 are empty in Type 2, so 800-386= 414 pokemon only have 1 Type, including Megas and 
different forms

~Question why are there 19 unique pokemon in Type 2?

```{r}
unique(pokemon$type_2)
```
"" is considered a type

Let's replace all "" in Type 2 with NAs

```{r}
pokemon$type_2 <- na_if(pokemon$type_2, "")

skim(pokemon)
```
This kept type_2 as a character type and replaced all the "" with NAs

What is the most common Type 1 and Type 2 groupings?

```{r}
pokemon %>% group_by(type_1, type_2) %>% 
  tally(sort=T) %>% 
  ungroup()
```

How many pokemon are there (including Megas and Forms)?
```{r}
pokemon %>% group_by(generation) %>% 
  tally(sort=T) %>% 
  ungroup()
```

Making a pivot table with stats all in one column, with the exception of total
```{r}
poke_pivot_long <- pokemon %>% pivot_longer(
  cols     = c('hp':'speed'),
  names_to = 'stats',
  values_to= 'y'
)

```

Plotting these values
```{r}
ggplot(data=poke_pivot_long) +
    geom_density(aes(x=y, group=stats, fill=stats), alpha=0.4)
```

Let's try but with boxplots now

```{r}
ggplot(data=poke_pivot_long) +
    geom_boxplot(aes(x=y, group=stats, fill=stats), 
                 alpha=0.4, 
                 outlier.colour="red",
                 outlier.fill="red",
                 outlier.size=3)
```

Reordering boxplots to match it with legend

```{r}
poke_pivot_long$stats <- factor(poke_pivot_long$stats,
                                 levels = c("speed", "sp_def", "sp_atk", "hp", "defense", "attack"))
```

played around with the parameters in aes 

```{r}
ggplot(data=poke_pivot_long) +
    geom_boxplot(aes(x=y, y=stats), 
                 alpha=0.4, 
                 outlier.colour="red",
                 outlier.fill="red",
                 outlier.size=3)
```
ok now attack is on top.. but they legend is out of order, so maybe I have to move the legend instead


Let's see how well pokemon stats correlate with each other

```{r}
correlation_stats <- pokemon %>% 
  select(hp, attack, defense, sp_atk, sp_def, speed) %>% 
  correlate()

correlation_stats
```

plotting correlation table
```{r}
rplot(correlation_stats)
```

Stats are weakly correlated with each other, some topping around 0.5 and as low as 0.1


single variable analysis

what is the most common pokemon type 1?

```{r}
pokemon %>% 
  group_by(type_1) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

what is the most common pokemon type 2?

```{r}
pokemon %>% 
  group_by(type_2) %>% 
  summarise(count = n()) %>% 
  top_n(n=10,wt = count) %>% 
  arrange(desc(count)) %>% 
  as.data.frame(na.rm=TRUE)
```

what is most cmmon type1-2?


difference total legendary vs non-legendary?

```{r}
pokemon %>% group_by(legendary) %>% 
    summarise(count = n(),
              hp = mean(hp),
              attack = mean(attack),
              defense = mean(defense),
              sp_atk = mean(sp_atk),
              sp_def = mean(sp_def),
              speed = mean(speed)
              ) 
```

Pretty telling that legendary pokemon are much stronger
