---
title: "pokemonAnalysisForFun"
author: "J"
date: "2024-12-27"
output: html_document
---

Just for fun analysis of Pokemon :)

```{r message=FALSE, warning=FALSE}
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable     # converting tables to HTML
)
```

Import data

```{r}
pokemon <- import("Pokemon.csv")
```

Exploratory Data Analysis

preview of the data
```{r}
skim(pokemon)
```

taking a look at the structure/type
```{r}
str(pokemon)
```

Cleaning Up Names (making it into lower case and switching space for underscore)

```{r}
pokemon <- pokemon %>% janitor::clean_names()

str(pokemon)
```

~Question how many pokemon have only 1 Type vs 2 Types?

Using Skimr, 386 are empty in Type 2, so 800-386= 414 pokemon only have 1 Type, including Megas and 
different forms

~Question why are there 19 unique pokemon in Type 2?

```{r}
unique(pokemon$type_2)
```
"" is considered a type

Let's replace all "" in Type 2 with NAs

```{r}
pokemon$type_2 <- na_if(pokemon$type_2, "")

skim(pokemon)
```
This kept type_2 as a character type and replaced all the "" with NAs

What is the most common Type 1 and Type 2 groupings?

```{r}
pokemon %>% group_by(type_1, type_2) %>% 
  tally(sort=T) %>% 
  ungroup()
```

How many pokemon are there (including Megas and Forms)?
```{r}
pokemon %>% group_by(generation) %>% 
  tally(sort=T) %>% 
  ungroup()
```

Making a pivot table with stats all in one column, with the exception of total
```{r}
poke_pivot_long <- pokemon %>% pivot_longer(
  cols     = c('hp':'speed'),
  names_to = 'stats',
  values_to= 'y'
)

```

Plotting these values
```{r}
ggplot(data=poke_pivot_long) +
    geom_density(aes(x=y, group=stats, fill=stats), alpha=0.4)
```

Let's try but with boxplots now

```{r}
ggplot(data=poke_pivot_long) +
    geom_boxplot(aes(x=y, group=stats, fill=stats), 
                 alpha=0.4, 
                 outlier.colour="red",
                 outlier.fill="red",
                 outlier.size=3)
```
That makes sense since Total is the addition of everything *facepalm*







Let's see how well pokemon stats correlate with each other

```{r}
correlation_stats <- pokemon %>% 
  select(hp, attack, defense, sp_atk, sp_def, speed) %>% 
  correlate()

correlation_stats
```

plotting correlation table
```{r}
rplot(correlation_stats)
```

Stats are weakly correlated with each other, some topping around 0.5 and as low as 0.1


ADDED NEW

try to compare difference between all the 6 stats
use normal vs normal, then normal vs leg, then within legendnardy



